% ----------------------------------------------------------------
% Letter ---------------------------------------------------------
% ----------------------------------------------------------------
\documentclass[12pt]{letter}
%\usepackage[spanish, activeacute]{babel}
\usepackage[cp1252]{inputenc}
% ----------------------------------------------------------------
\signature{Fernando Mazzone}


\begin{document}
\hyphenation{si-gui-en-tes}
% ----------------------------------------------------------------
\begin{letter}{ % Addressee
  }
 
\opening{ Dear Editor-in-Chief Mr. Mitidieri:}

We would like to submit the revised version of our manuscript entitled  ``Some existence results on periodic solutions of Euler-Lagrange equations in an Orlicz-Sobolev space setting'', Ms. Ref. No.: NA-D-14-01591, for consideration by the ``Nonlinear Analysis: Theory, Methods \& Applications ''.
Next, we enumerate the list of changes according to the reviewers' comments.

\begin{itemize}

\item Page 2: we added the spaces to which the functions $b$, $c$, and $f$ belong.
\item Page 4: we added a reference to the H\"older inequality in Orlicz spaces.
\item Page 5:  we deleted the requirement that $u$ belongs to $L^{\Phi}$  in the definition of the Orlicz-Sobolev space $W^1L^{\Phi}$.
\item Page 6: we renamed inequalities (15) and (16) and we corrected the references  to them in whole paper.
\item Page 9, Definition 3.1: we changed the definition of  Carath\'eodory function according to reviewers' advise.
\item Page 17, Lemma 6.1: We added  the hypothesis ``F satisfies the condition (A)'' because the differentiability of F is necessary and we were not assuming that condition in the previous formulation.
\item Page 18: before Theorem 6.2, we added a brief comment.  
\item Pages 19 and 20: after the proof of Theorem 6.2, we included a discussion about the contributions of our main theorem and examples of its applications. 

\end{itemize}
We remain at your disposal for any further information you may need.


\closing{Sincerely, }
\end{letter}
% ----------------------------------------------------------------
\end{document}
% ----------------------------------------------------------------
