
\documentclass[twoside]{article}


\NeedsTeXFormat{LaTeX2e}
\ProvidesPackage{mathscinet}[2002/04/17 v1.05]
\RequirePackage{textcmds}\relax
\ProvideTextCommandDefault{\cprime}{\tprime}



%\usepackage{hyperref}
\usepackage{amssymb,amsthm}
\usepackage{amsmath}
\usepackage{color}
\usepackage{ esint }

\usepackage{fancyhdr}
\usepackage{times}

\usepackage[latin1]{inputenc}

\usepackage{comment}
\usepackage{url}
\usepackage{xcolor}
\usepackage{adjustbox}
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}

\newtheorem{defi}[thm]{Definition}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{remark}
\newtheorem{comentario}{Remark}


\title{Periodic solutions of 
Euler-Lagrange equations with ``sublinear nonlinearity'' in an Orlicz-Sobolev space setting}
\author{Sonia Acinas \thanks{SECyT-UNRC, UNSL and CONICET}\\
Instituto de Matem\'atica Aplicada San Luis (CONICET-UNSL)\\
(5700) San Luis, Argentina\\
Universidad Nacional de La Pampa\\
(6300) Santa Rosa, La Pampa, Argentina\\
\url{sonia.acinas@gmail.com}\\[3mm]
Fernando D. Mazzone \thanks{SECyT-UNRC and CONICET}\\
Dpto. de Matem\'atica, Facultad de Ciencias Exactas, F\'{\i}sico-Qu\'{\i}micas y Naturales\\
Universidad Nacional de R\'{i}o Cuarto\\
(5800) R\'{\i}o Cuarto, C\'ordoba, Argentina,\\
\url{fmazzone@exa.unrc.edu.ar}
}

\date{}

\newcommand{\orlnor}{\|_{L^{\Phi}}}
\newcommand{\lurnor}{\|^{*}_{L^{\Phi}}}
\newcommand{\linf}{\|_{L^{\infty}}}
\newcommand{\lphi}{L^{\Phi}}
\newcommand{\lpsi}{L^{\Psi}}
\newcommand{\ephi}{E^{\Phi}}
\newcommand{\claseor}{C^{\Phi}}
\newcommand{\wphi}{W^{1}\lphi}
\newcommand{\sobnor}{\|_{W^{1}\lphi}}
\newcommand{\domi}{\mathcal{E}^{\Phi}_d(\lambda)}
\renewcommand{\b}[1]{\boldsymbol{#1}}
\newcommand{\rr}{\mathbb{R}}
\newcommand{\nn}{\mathbb{N}}
\newcommand{\ccdot}{\b{\cdot}}
\renewcommand{\leq}{\leqslant} 
\newcommand{\epsi}{E^{\Psi}}

\begin{document}



\maketitle
%
\begingroup%Locallizing the change to `thefootnote'.
    \renewcommand{\thefootnote}{}%Removing the footnote symbol.
    %
    \footnotetext{%
    %   2010 Mathematics Subject Classification
    %   http://www.ams.org/msc/
    \textbf{2010  AMS Subject Classification.} Primary: .
    Secondary: .
    }%
        \footnotetext{%
    \textbf{Keywords and phrases.}  .
    }%
    \endgroup
%
%
%
%

\begin{abstract}
In this paper we....

ALGO AS\'I HAB\'IA QUE ESCRIBIR AC\'A O EN LA INTRO....

The results of this paper improve on the classic ones obtained in \cite{tang1998periodic} and \cite{tang2010periodic} with $|\nabla F(t,\b{x})|\leq f(t) \b{x}^{\alpha}+g(t)$
for $\alpha<1$ and $\alpha<p$ respectively. Here, the bounds of the sublinearity are bigger than those considered in the aforementioned works.
\end{abstract}




\pagestyle{fancy} \headheight 35pt \fancyhead{} \fancyfoot{}

\fancyfoot[C]{\thepage} \fancyhead[CE]{\nouppercase{S. Acinas and F.D. Mazzone }} \fancyhead[CO]{\nouppercase{\section}}

\fancyhead[CO]{\nouppercase{\leftmark}}


%\tableofcontents




\section{Introduction}
This paper is concerned with the existence of periodic solutions of the problem
\begin{equation}\label{ProbPrin}
    \left\{%
\begin{array}{ll}
   \frac{d}{dt} D_{y}\mathcal{L}(t,\b{u}(t),\b{\dot{u}}(t))= D_{\b{x}}\mathcal{L}(t,\b{u}(t),\b{\dot{u}}(t)) \quad \hbox{a.e.}\ t \in (0,T)\\
    \b{u}(0)-\b{u}(T)=\b{\dot{u}}(0)-\b{\dot{u}}(T)=0
\end{array}%
\right.
\end{equation}
where $T>0$, $\b{u}:[0,T]\to\rr^d$ is absolutely continuous and the \emph{Lagrangian} $\mathcal{L}:[0,T]\times\rr^d\times\rr^d\to\rr$ is a Carath\'eodory function satisfying the conditions
\begin{eqnarray}
|\mathcal{L}(t,\b{x},\b{y})| &\leq a(|\b{x}|)\left(b(t)+ \Phi\left(\frac{|\b{y}|}{\lambda}+f(t) \right)\right),\label{cotaL}\\
|D_{\b{x}}\mathcal{L}(t,\b{x},\b{y})| &\leq a(|\b{x}|)\left(b(t)+ \Phi\left(\frac{|\b{y}|}{\lambda}+f(t) \right)\right),\label{cotaDxL}\\
|D_{\b{y}}\mathcal{L}(t,\b{x},\b{y})| &\leq a(|\b{x}|)\left(c(t)+ \varphi\left(\frac{|\b{y}|}{\lambda}+f(t)\right)  \right).\label{cotaDyL}
\end{eqnarray}
In these inequalities we assume that  $a\in C(\mathbb{R}^+,\mathbb{R}^+)$, $\lambda>0$, $\Phi$ is an $N$-function (see section  Preliminaries  for definitions), $\varphi$ is the right continuous derivative of $\Phi$. The non negative functions $b,c$ and $f$ satisfy that  $b\in L^1_1([0,T]) $,  $c\in\lpsi_1([0,T])$ and  $f\in \ephi_1([0,T])$, where  the Banach spaces $ L^1_1([0,T]), \lpsi_1([0,T])$ and  $\ephi_1([0,T])$  will be defined later.


It is well known that problem \eqref{ProbPrin} comes from a variational one, that is,  a solution of \eqref{ProbPrin}  
is a critical point of the \emph{action integral}
\begin{equation}\label{integral_accion}
I(\b{u})=\int_{0}^T \mathcal{L}(t,\b{u}(t),\b{\dot{u}}(t))\ dt.
\end{equation}




\section{Preliminaries}\label{preliminares}

For reader convenience, we give a short introduction to Orlicz and Orlicz-Sobolev spaces of vector valued functions and a  list  of results that we will use throughout the article. 
Classic references for Orlicz spaces of real valued functions are \cite{adams_sobolev,KR,rao1991theory}.
For  Orlicz spaces of vector valued functions, see \cite{Orliczvectorial2005} and the references therein.

Hereafter we denote  by $\mathbb{R}^+$  the set of all non negative real numbers. A function $\Phi:\mathbb{R}^+\to \mathbb{R}^+ $ is called an \emph{$N$-function} if $\Phi$ is given by 
\[
\Phi(t)=\int_{0}^t \varphi(\tau)\ d\tau,\quad\hbox{for } t\geq 0,
\]
where $\varphi:\mathbb{R}^+\rightarrow \mathbb{R}^+$ is a right continuous non decreasing function  satisfying   $\varphi(0)=0$, $\varphi(t)>0$ for $t>0$ and
$\lim_{t\rightarrow \infty}\varphi(t)=+\infty$.

Given a function $\varphi$ as above, we  consider the so-called right inverse function $\psi$ of $\varphi$ which is 
defined by $\psi(s)=\sup_{\varphi(t)\leq s}t$.
The function $\psi$ satisfies the same properties as the function $\varphi$, therefore we have an $N$-function $\Psi$ such that $\Psi'=\psi$ .
 The function $\Psi$ is called the \emph{complementary function} of $\Phi$.


We say that $\Phi$ satisfies the  \emph{$\Delta_2$-condition}, denoted by $\Phi \in \Delta_2$, 
if there exist  constants $K>0$ and  $t_0\geq 0$ such that 
\begin{equation}\label{delta2defi}\Phi(2t)\leq K\Phi(t)
\end{equation}
for every $t\geq t_0$. 
If $t_0=0$,  we say that $\Phi$ satisfies the \emph{$\Delta_2$-condition globally} ($\Phi \in \Delta_2$ globally).  

% and plain symbols indicate scalars.

Let $d$ be a positive integer. We denote by $\mathcal{M}_d:=\mathcal{M}_d([0,T])$ the set of all measurable functions defined on $[0,T]$ with values on $\mathbb{R}^d$ and  we write $\b{u}=(u_1,\dots,u_d)$ for  $\b{u}\in \mathcal{M}_d$.
In this paper we adopt the convention that bold symbols denote points in $\mathbb{R}^d$.


Given  an $N$-function $\Phi$ we define the \emph{modular function} 
$\rho_{\Phi}:\mathcal{M}_d\to \mathbb{R}^+\cup\{+\infty\}$ by
\[\rho_{\Phi}(\b{u}):= \int_0^T \Phi(|\b{u}|)\ dt.\]
Here $|\cdot|$ is the euclidean norm of $\mathbb{R}^d$.
The \emph{Orlicz class} $C_d^{\Phi}=C_d^{\Phi}([0,T])$  is given  by
\begin{equation}\label{claseOrlicz}
  C^{\Phi}_d:=\left\{\b{u}\in \mathcal{M}_d | \rho_{\Phi}(\b{u})< \infty \right\}.
\end{equation}
The \emph{Orlicz space} $\lphi_d=L^{\Phi}_d([0,T])$ is the linear hull of $\claseor_d$;
equivalently,
\begin{equation}\label{espacioOrlicz}
\lphi_d:=\left\{ \b{u}\in \mathcal{M}_d | \exists \lambda>0: \rho_{\Phi}(\lambda \b{u}) < \infty   \right\}.
\end{equation}
  The Orlicz space $\lphi_d$ equipped with the \emph{Orlicz norm}
\[
\|  \b{u}  \orlnor:=\sup \left\{  \int_0^T \b{u}\b{\cdot} \b{v}\ dt \big| \rho_{\Psi}(\b{v})\leq 1\right\},
\]
is a Banach space. By $\b{u}\b{\cdot} \b{v}$ we denote the usual dot product in $\mathbb{R}^{d}$ between $\b{u}$ and $\b{v}$.  
The following alternative expression for the norm, known as \emph{Amemiya norm},     will  be useful (see \cite[Thm. 10.5]{KR} and \cite{hudzik2000amemiya}). For every $\b{u}\in\lphi$,

\begin{equation}\label{amemiya}
\|\b{u}\orlnor=\inf\limits_{k>0}\frac{1}{k}\left\{1+\rho_{\Phi}(k\b{u})\right\}.
\end{equation}



The subspace $\ephi_d=\ephi_d([0,T])$ is defined as the closure in $\lphi_d$ of the subspace $L^{\infty}_d$ of all $\mathbb{R}^d$-valued essentially bounded functions. It is shown that  $\ephi_d$ is the only one maximal subspace contained in the Orlicz class $\claseor_d$, i.e. 
$\b{u}\in\ephi_d$ if and only if $\rho_{\Phi}(\lambda \b{u})<\infty$ for any $\lambda>0$.  

A generalized version of \emph{H\"older's inequality} holds in Orlicz spaces (see \cite[Th. 9.3]{KR}). Namely, if $\b{u}\in\lphi_d$ and $\b{v}\in\lpsi_d$ then $\b{u}\ccdot\b{v}\in L_1^1$ and
\begin{equation}\label{holder}
\int_0^T\b{v}\ccdot\b{u}\ dt\leq \|\b{u}\orlnor\|\b{v}\|_{L^{\Psi}}.
\end{equation}




If $X$ and $Y$ are  Banach spaces such that  $Y\subset X^*$, we denote by $\langle\cdot,\cdot\rangle:Y\times X\to\mathbb{R}$ the bilinear pairing  map given by $\langle x^*,x\rangle=x^*(x)$. H\"older's inequality shows that $\lpsi_d\subset \left[\lphi_d\right]^*$, where the pairing  
$\langle \b{v}, \b{u}\rangle$
is defined by 
\begin{equation}\label{pairing}
  \langle \b{v},\b{u}\rangle=\int_0^T\b{v}\ccdot\b{u}\ dt
\end{equation}
with  $\b{u}\in\lphi_d$ and $\b{v}\in\lpsi_d$.
 Unless $\Phi \in \Delta_2$, the relation $\lpsi_d= \left[\lphi_d\right]^*$ will not hold. In general, it is true  that  $\left[\ephi_d\right]^*=\lpsi_d$.


Like in \cite{KR}, we will consider the subset $\Pi(\ephi_d,r)$ of $\lphi_d$ given by
\[\Pi(\ephi_d,r):=\{\b{u}\in\lphi_d| d(\b{u},\ephi_d)<r\}.\]
This set is related to the Orlicz class $\claseor_d$ by means of inclusions, namely,
\begin{equation}\label{inclusiones}\Pi(\ephi_d, r )\subset r \claseor_d\subset\overline{\Pi(\ephi_d,r)}
\end{equation}
for any positive $r$.
If $\Phi \in \Delta_2$,  then the sets $\lphi_d$, $\ephi_d$, $\Pi(\ephi_d,r)$ and $\claseor_d$ are equal.



We define the \emph{Sobolev-Orlicz space} $\wphi_d$ (see \cite{adams_sobolev}) by
\[\wphi_d:=\{\b{u}| \b{u} \hbox{ is absolutely continuous and } \b{\dot{u}}\in \lphi_d\}.\]
$\wphi_d$ is a Banach space when equipped with the norm
\begin{equation}\label{def-norma-orlicz-sob}
\|  \b{u}  \|_{\wphi}= \|  \b{u}  \|_{\lphi} + \|\b{\dot{u}}\orlnor.
\end{equation}



For a  function $\b{u}\in L^1_d([0,T])$, we write $\b{u}=\overline{\b{u}}+\widetilde{\b{u}}$ where $\overline{\b{u}} =\frac1T\int_0^T \b{u}(t)\ dt$ and $\widetilde{\b{u}}=\b{u}-\overline{\b{u}}$.

As usual, if $(X,\|\cdot\|_X)$ is a Banach space and $(Y,\|\cdot \|_Y)$ is a subspace of $X$,  we write $Y\hookrightarrow X$ and we say that $Y$ is \emph{embedded} in $X$  when the restricted identity map $i_Y:Y\to X$ is bounded. That is, there exists $C>0$ such that  for any $y\in Y$ we have $\|y\|_X\leq C\|y\|_Y$.  With this notation, H\"older's inequality states that  $\lpsi_d\hookrightarrow  \left[\lphi_d\right]^*$; and, it is easy to see that for every $N$-function $\Phi$ we have that $L^{\infty}_d\hookrightarrow\lphi_d \hookrightarrow L^1_d$.


 Recall that a function   $w:\mathbb{R}^+\to \mathbb{R}^+$ is called  a \emph{modulus of continuity} if $w$ is a continuous increasing function which satisfies $w(0)=0$. For example, it can be easily shown that $w(s)=s\Phi^{-1}(1/s)$ is a modulus of  continuity for every $N$-function $\Phi$.  We say that $\b{u}:[0,T]\to\rr^d$  has modulus of continuity $w$  when there exists a constant $C>0$ such that 
\begin{equation}\label{w-holder}|\b{u}(t)-\b{u}(s)|\leq Cw(|t-s|).
\end{equation}


We denote by $C^w([0,T],\rr^d)$  the space of  $w$-H\"older continuous functions. This is the space of all functions satisfying \eqref{w-holder} for some $C>0$ and it is a Banach space with norm
\[\|\b{u}\|_{  C^w([0,T],\rr^d) }  :=\|\b{u}\|_{L^{\infty}}+\sup\limits_{t\neq s}\frac{|\b{u}(t)-\b{u}(s)|}{w(|t-s|)}.\]





 An important aspect of the theory of Sobolev spaces is related to embedding theorems. There is an extensive literature on this question in the  Orlicz-Sobolev space setting, see for example
 \cite{cianchi2000fully,cianchi1999some,claverooptimal,edmunds2000optimal,kerman2006optimal}.
The next simple lemma, whose proof can be found in \cite{ABGMS2015}, will be used systematically.




\begin{lem}\label{inclusion orlicz} Let  $w(s):= s\Phi^{-1}(1/s)$. Then, the following statements hold:
\begin{enumerate}
\item\label{inclusion orlicz_item1} $\wphi\hookrightarrow C^w([0,T],\rr^d) $ and for every $\b{u}\in\wphi$
\begin{align}
 &\left|\b{u}(t)-\b{u}(s) \right| \leq  \|\b{\dot{u}}\orlnor w(| t-s|),&\label{in-sob-cont}
\\
& \|\b{u}\|_{L^{\infty}} \leq\Phi^{-1}\left(\frac{1}{T}\right)\max\{1,T\}\|\b{u}\sobnor&\label{sobolev}
\end{align}
\item For every $\b{u}\in\wphi$ we have $\widetilde{\b{u}}\in L^{\infty}_d$ and 
\begin{align}
& \|\widetilde{\b{u}}\|_{L^{\infty}} \leq T\Phi^{-1}\left(\frac{1}{T}\right)\|\b{\dot u}\orlnor&\text{  (Sobolev's inequality).}\label{wirtinger}
\end{align}




\end{enumerate}
\end{lem}


The following result is analogous to some lemmata in $W^1L^p_d$, see \cite{xu2007some}.
\begin{lem}\label{infinito-a-prom-upunto}
If $\|\b{u}\sobnor\to \infty$, then $(|\b{\overline u}|+\|\b{\dot u}\orlnor)\to \infty$.
\end{lem}

\begin{proof}
By the decomposition $\b{u}=\b{\overline u}+\b{\tilde{u}}$ and some elementary operations, 
we get
\begin{equation}\label{cota-u-lphi}
\|\b{u}\orlnor=
\|\b{\overline u}+\b{\tilde{u}}\orlnor\leq 
\|\b{\overline u}\orlnor+\|\b{\tilde{u}}\orlnor=
|\b{\overline u}|\|1\orlnor+\|\b{\tilde{u}}\orlnor.
\end{equation}
It is known that $L^{\infty}_d\hookrightarrow\lphi_d$, i.e.,
there exists $C_1=C_1(T)>0$ such that for any $\b{\tilde{u}}\in L^{\infty}_d$ we have
\[
\|\b{\tilde{u}}\orlnor
\leq 
C_1 \|\b{\tilde{u}}\|_{L^{\infty}};
\]
and, applying  Sobolev's inequality,  we obtain the Wirtinger's inequality, that is there exists $C_2=C_2(T)>0$ suvh that 
\begin{equation}\label{cota-u-tilde}
\|\b{\tilde{u}}\orlnor
\leq 
C_2\|\b{\dot{u}}\orlnor.
\end{equation}

Therefore, from \eqref{cota-u-lphi}, \eqref{cota-u-tilde} and \eqref{def-norma-orlicz-sob}, 
we get
\[
\|\b{u}\sobnor\leq
C_3(|\b{\overline u}|+\|\b{\dot{u}}\orlnor)
\]
where $C_3=C_3(T)$. Finally, as $\|\b{u}\sobnor\to \infty$ we conclude that   
$(|\b{\overline u}|+\|\b{\dot{u}}\orlnor)\to \infty$.
\end{proof}


We present a definition that will be useful later.
 
\begin{defi} A function $\mathcal{L}:[0,T]\times \mathbb{R}^d \times \mathbb{R}^d \rightarrow \mathbb{R}$ is a \emph{Carath\'eodory} function if for fixed $(\b{x},\b{y})$
the map $t \mapsto \mathcal{L}(t, \b{x},\b{y})$ is measurable  and for fixed $t$ the map  $(\b{x},\b{y}) \mapsto \mathcal{L}(t, \b{x}, \b{y})$ is continuous  for almost everywhere $t\in [0,T]$. We say that 
$\mathcal{L}(t, \b{x},\b{y})$ is  \emph{differentiable Carath\'eodory} if in addition $\mathcal{L}(t, \b{x},\b{y})$ is continuously differentiable with respect to $\b{x}$ and $\b{y}$  for almost everywhere $t\in [0,T]$.

\end{defi}


In \cite{ABGMS2015} we proved the next results.

\begin{thm}\label{teorema_acotacion}
Let $\mathcal{L}$ be a differentiable Carath\'eodory function satisfying \eqref{cotaL}, \eqref{cotaDxL} and \eqref{cotaDyL}. 
Then the following statements hold:
\begin{enumerate}
\item \label{T1item1} \label{A1} The action integral given by \eqref{integral_accion}
is finitely defined on $\domi:=W^{1}\lphi_d\cap\{\b{u}|\b{\dot{u}}\in\Pi(\ephi_d,\lambda)\}$.

\item\label{T1item3} The function  $I$ is G\^ateaux differentiable on $\domi$ and  its derivative $I'$ is demicontinuous from $\domi$  into $\left[\wphi_d \right]^*$. Moreover, $I'$ is given by the following expression
\begin{equation}\label{DerAccion}
\langle  I'(\b{u}),\b{v}\rangle= \int_0^T \left\{D_{\b{x}}\mathcal{L}\big(t,\b{u},\b{\dot{u}}\big)\ccdot \b{v}+ D_{\b{y}}\mathcal{L}\big(t,\b{u},\b{\dot{u}}\big)\ccdot\b{\dot{v}}\right\} \ dt.
\end{equation}

\item\label{T1item4}  If  $\Psi \in \Delta_2$ then 
  $I'$ is continuous from $\domi$ into $\left[\wphi_d\right]^*$ when both spaces are equipped with the strong topology.
\end{enumerate}
\end{thm}





In \cite{ABGMS2015} we derive the Euler-Lagrange equations associated to critical points of action integrals on the subspace of $T$-periodic functions.  
We denote by $\wphi_T$ the subspace of $\wphi_d$ containing all  $T$-periodic functions. As usual, when $Y$ is a subspace of
the Banach space $X$, we denote by $Y^{\perp}$ the \emph{annihilator subspace} of $X^*$, i.e. the subspace
that consists of all  bounded linear functions which are identically zero on $Y$.

We recall that  a function $f: \mathbb{R}^d \to \mathbb{R}$ is called \emph{strictly convex} if 
$f\left(\tfrac{\b{x}+\b{y}}{2}\right)< \tfrac{1}{2} \left(f\left(
\b{x}\right)+f\left( \b{y}\right)\right)$ for  $\b{x}\neq\b{y}$.  
It is  well known that if $f$ is a strictly convex and differentiable function, then
$D_{\b{x}}f:\mathbb{R}^d\to\mathbb{R}^d$ is a one-to-one map  (see, e.g. \cite[Thm. 12.17]{rockafellar2009variational}).


\begin{thm}\label{critpoint} Let $\b{u}\in\domi$ be  a $T$-periodic function. The following statements are equivalent:
\begin{enumerate}
 \item $I'(\b{u})\in\left( \wphi_T\right)^{\perp}$.
 \item  $D_{\b{y}}\mathcal{L}(t,\b{u}(t),\b{\dot{u}}(t))$ is an absolutely continuous function and $\b{u}$ solves the following boundary value problem
 \begin{equation}\label{ecualagran2}
    \left\{%
\begin{array}{ll}
   \frac{d}{dt} D_{y}\mathcal{L}(t,\b{u}(t),\b{\dot{u}}(t))= D_{\b{x}}\mathcal{L}(t,\b{u}(t),\b{\dot{u}}(t)) \quad \hbox{a.e.}\ t \in (0,T)\\
    \b{u}(0)-\b{u}(T)=D_{\b{y}}\mathcal{L}(0,\b{u}(0),\b{\dot{u}}(0))-D_{\b{y}}\mathcal{L}(T,\b{u}(T),\b{\dot{u}}(T))=0.
\end{array}%
\right.
\end{equation}
\end{enumerate}
Moreover if $D_{\b{y}}\mathcal{L}(t,x,y)$ is $T$-periodic with respect to the variable $t$ and strictly convex with respect to $\b{y}$, then
$D_{\b{y}}\mathcal{L}(0,\b{u}(0),\b{\b{\dot{\b{u}}}}(0))-D_{\b{y}}\mathcal{L}(T,\b{u}(T),\b{\dot{u}}(T))=0$ is equivalent to $\b{\dot{u}}(0)=\b{\dot{u}}(T)$.
\end{thm}


{\bf Habr\'ia que ver si el lugar de los \'indices es el adecuado. Copi\'e lo que ten\'iamos en el 
primer trabajo.}


Next, we enumerate some definitions and results from the theory of convex functions. 
We suggest \cite{FK97, GP77, KR, M, rao1991theory} for definitions, proofs and additional details.

We denote by $\alpha_{\varphi}$ and $\beta_{\varphi}$ the so called  \emph{Matuszewska-Orlicz indices} of the function $\varphi$, which are defined next. Given
an increasing, unbounded, continuous function  \linebreak $\varphi:[ 0,+\infty)\to [0,+\infty)$ such that $\varphi(0)=0$ we define
\begin{equation}\label{MO_indices}
    \alpha_{\varphi}:=\lim\limits_{t\to 0^{+}}\frac{\log \left (\sup\limits_{u>0}\frac{\varphi(t u)}{\varphi(u)} \right ) }{\log(t)},\quad
    \beta_{\varphi}:=\lim\limits_{t\to +\infty}\frac{\log \left  (\sup\limits_{u>0}\frac{\varphi(t u)}{\varphi(u)}\right )}{\log(t)}.
\end{equation}
We have that $0\leq \alpha_{\varphi}\leq \beta_{\varphi}\leq +\infty$. The relation $\beta_{\varphi}<\infty$ holds true if and only if $\varphi$ is a
$\Delta_2$-function. If $\varphi$ is a homeomorphism  we have that
\begin{equation}\label{inv_indices}
    \alpha_{\varphi^{-1}}=\frac{1}{\beta_{\varphi}}.
\end{equation}
Moreover $\varphi\in\mathcal{F}$ implies  $\alpha_{\varphi}\geq 1$. As a consequence, $\varphi^{-1}$ is a $\Delta_2$-function.

 It is well known   that if $\varphi$ is an increasing $\Delta_2$-function, $\varphi$ is controlled by above and below 
 by power functions.  More concretely, for every $\epsilon>0$ there exists a
constant $K=K(\varphi,\epsilon)$ such that, for every $t,u\geq 0$,
\begin{equation}\label{delta2-potencias}
    K^{-1}\min\big\{t^{\beta_{\varphi}+\epsilon},t^{\alpha_{\varphi}-\epsilon} \big\}\varphi(u)\leq \varphi(t u)\leq
    K\max\big\{t^{\beta_{\varphi}+\epsilon},t^{\alpha_{\varphi}-\epsilon} \big\}\varphi(u).
\end{equation}



\begin{lem}
Let $\Phi,\Psi$ complementary functions.
The next statements are equivalent:
\begin{enumerate}
\item\label{item1} $\Psi \in \Delta_2$ globally.
\item\label{item2} There exists an $N$-function $\Phi_1 \in \Delta_2$ such that
\begin{equation}\label{eq:caract_delta2}
\Phi(rs)\geq \Phi_1(r)\Phi(s)\;\;\mbox{for every}\;\;r\geq1,\;\;s\geq 0.
\end{equation}
\end{enumerate}
\end{lem}

\begin{proof}
\ref{item1})$\Rightarrow$\ref{item2}) As $\Psi \in  \Delta_2$ globally, there exist $k>0$ and $\nu>1$ such that
\[
\Phi(rs)\geq k r^\nu \Phi(s)\;\;r\geq 1,\;s>0,
\]
which is \eqref{eq:caract_delta2} with $\Phi_1(r)=kr^\nu$ that is an $N$-function that satisfies the $\Delta_2$-condition.

\ref{item2})$\Rightarrow$\ref{item1})
Next, we follow  \cite[p. 32, Prop. 13]{rao1991theory} and \cite[p. 29, Prop. 9]{rao1991theory}.
Assume that 
\[
\Phi_1(r)\Phi(s)\leq \Phi(rs)\;\;r>1,\;s\geq 0.
\]
Let $u=\Phi_1(r)\geq \Phi_1(1)$ and $v=\Phi(s)\geq 0$. By the well known inequality \cite[p. 13, Prop. 1]{rao1991theory} and \eqref{eq:caract_delta2} we have  for $u\geq \Phi_1(1)$ and $v\geq 0$
\[
\frac{uv}{\Psi^{-1}(uv)}\leq \Phi^{-1}(uv)\leq\Phi_1^{-1}(u)\Phi^{-1}(v)\leq
\frac{4uv}{\Psi_1^{-1}(u)\Psi^{-1}(v)},
\]
then 
\[
\Psi^{-1}_1(u)\Psi^{-1}(v)\leq 4 \Psi^{-1}(uv).
\]
If we take $x=\Psi^{-1}_1(u)\geq \Psi^{-1}_1(\Phi_1(1))$ and $y=\Psi^{-1}(v)\geq 0$, then 
\[
\Psi\left(\frac{xy}{4}\right)\leq \Psi_1(x)\Psi(y).
\]
Now, taking  $x\geq \max\{8,\Psi_1^{-1}(\Phi_1(1))\}$ we get that $\Psi \in \Delta_2$ globally.
\end{proof}

The following Lemma generalizes \cite[Lemma 5.2]{ABGMS2015}.

\begin{lem}\label{lem_coer}
Let $\Phi$ be $N$-functions and suppose that $\Psi \in \Delta_2$ globally. Then
\begin{equation}\label{eq:coer_mod}
\lim\limits_{\|u\orlnor\to \infty}
\frac{\int_0^T \Phi(|u|)\,dt}{\Phi_0(\|u\orlnor)}=\infty,
\end{equation}


for every $\Phi_0$ with $\Phi_0=o(\Phi_1)$ at $\infty$ where $\Phi_1$ is any $N$-function satisfying \eqref{eq:caract_delta2}.

Reciprocally if  \eqref{eq:coer_mod} holds for some $N$-function $\Phi_0$ then $\Psi\in\Delta_2$ (at $\infty$). 
\end{lem}

\begin{proof}
By the assumptions on $\Phi$ and $\Phi_1$  and identity \eqref{amemiya}, we have
\[
\frac{\int_0^T \Phi(|u|)\,dt}{\Phi_0(\|u\orlnor)}\geq
\Phi_1(r) \frac{\int_0^T \Phi(r^{-1}|u|)\,dt}{\Phi_0(\|u\orlnor)}\geq
\frac{\Phi_1(r)}{\Phi_0(\|u\orlnor)}\{r^{-1}\|u\orlnor-1\}.
\]
Now, we choose $r=\frac{\|u\orlnor}{2}$, as $\|u\orlnor\to\infty$ we can assume $r>1$, we use the fact that $\Phi_1\in\Delta_2$ and
$\Phi_0=o(\Phi_1)$ at $\infty$, and 
therefore we get
\[
\lim\limits_{\|u\orlnor \to \infty} \frac{\int_0^T \Phi(|u|)\,dt}{\Phi_0(\|u\orlnor)}\geq
\lim\limits_{\|u\orlnor \to \infty} \frac{\Phi_1\left(\frac{\|u\orlnor}{2}\right)}{\Phi_0(\|u\orlnor)}
\geq
C \lim\limits_{\|u\orlnor \to \infty} \frac{\Phi_1(\|u\orlnor)}{\Phi_0(\|u\orlnor)}=\infty.
\]
The last affirmation of the Lemma follows from the fact that if $\Phi_0$ is $N$-function then $\Phi_0(u)\geq ku$ for  $k$ small enough. Therefore \eqref{eq:coer_mod} holds for $\Phi_0(u)=|u|$, then \cite[Lemma 5.2]{ABGMS2015}  implies  $\Psi\in\Delta_2$ at $\infty$.

\end{proof}


\begin{comentario}  We point out that this Lemma can be applied to more cases than \cite[Lemma 5.2]{ABGMS2015}. For example, if $\Phi(u)=u^2$, $\Phi_1$ and $\Phi_0$ are a $N$-functions with principal parts equal to $u^2/\log u$ and $u^2/(\log u)^2$ respectively (see \cite[p. 16]{KR} and \cite[Section 7]{KR} for the definition and properties of principal part). Then  \eqref{eq:coer_mod} holds for $\Phi_0$, however $\Phi_0(u)$ is not dominated for any  power function $|u|^{\alpha}$ for every $\alpha<2$. 
\end{comentario}






We define the following  functionals $J_{C,\Phi_0}:\lphi\to (-\infty,+\infty]$ and $  H_{C,\Phi_0}:\rr^n\to \rr$, with $C>0$ and $\Phi_0$ $N$-function, by
\begin{equation}\label{func_phi}
  J_{C,\Phi_0}(\b{u}):= \rho_{\Phi}\left(\b{u}\right)-C\Phi_0\left(\|\b{u}\orlnor\right),
\end{equation}
 and

\begin{equation}\label{eq:functional_H-bis}
 H_{C,\Phi_0}(\b{x}):=\int_0^TF(t,\b{x})dt-C\Phi_0(|\b{x}|),
\end{equation}
respectively.





\section{Lagrangians satisfying  sublinear nonlinearity type conditions}


Like in \cite{ABGMS2015} we consider Lagrangians $\mathcal{L}$ which are lower bounded as follows 
\begin{equation}\label{cota_inf}
\mathcal{L}(t,\b{x},\b{y})\geq \alpha_0\Phi\left(\frac{|\b{y}|}{\Lambda}\right)+ F(t,\b{x}).
\end{equation}

If $\mathcal{L}$ is given by right hand side in \eqref{cota_inf} and $\Phi(u)=|u|^2$ then the ODE $\ddot{\b{u}}=\nabla F(t,\b{u}(t))$ in \eqref{ProbPrin}  is quasilinear,  being  $\nabla F(t,\b{u}(t))$ the nonlinearity. Following the literature, we refer to $\nabla F$ as the non linearity yet we assume in \eqref{cota_inf} just the inequality. In \cite{tang1998periodic} and \cite{tang2010periodic} the authors considered for $p$-laplacian non linearities satisfying the inequality
\[ |\nabla F(t,\b{x})|\leq b_1(t)|\b{x}|^{\alpha}+b_2(t),\]
where  $b_1,b_2 \in L^1_1$ and $\alpha$ is any power less than $p$. Thus they said $F$ is a sublinear nonlinearity. In this paper we consider the following type of bounds for the nonlinearity $F$
\begin{equation}\label{holder_cont-mu}
  \left| \nabla F(t,\b{x}) \right|\leq b_1(t)\varphi_0(|\b{x}|)+b_2(t),
\end{equation}
where $\varphi_0=\Phi'_0$ with $\Phi_0$ an $N$-function. The employment of  $N$-functions instead of power functions in  inequalities likes  \eqref{holder_cont-mu}  will allow us to extend some results of   \cite{tang1998periodic} and \cite{tang2010periodic} even in the $p$-Laplacian case.


Based on \cite{mawhin2010critical} we say that $F$ satisfies the condition (A) if  $F(t,\b{x})$ is a Carath\'eo\-dory function and  $F$ is continuously differentiable with respect to $\b{x}$. Moreover, the next inequality holds 
\begin{equation}\label{condA2}|F(t,\b{x})|+ |D_{\b{x}}F(t,\b{x})|\leq a(|\b{x}|)b_0(t),\quad\text{for a.e. }t\in [0,T], \forall\b{x}\in\rr^d.
\end{equation}

The following theorem establishes  coercivity of $I$ assuming sublinear conditions on the nonlinearity  $\nabla F$. 












\begin{thm}\label{coercitividad-r}
Let  $\mathcal{L}$ be a lagrangian function satisfying \eqref{cotaL}, \eqref{cotaDxL}, \eqref{cotaDyL}, \eqref{cota_inf}  and $F$ satisfies condition (A). We assume the following conditions:
\begin{enumerate}
\item $\Psi\in\Delta_2$.
\item Inequality \eqref{holder_cont-mu} with $b_1,b_2 \in L^1_1$,  $\varphi_0=\Phi'_0$, $\Phi_0$ differentiable and  $\Delta_2$ globally $N$-function, 
$\Phi_0=o(\Phi_1)$ at $\infty$, where $\Phi_1$ satisfies \eqref{eq:caract_delta2}.
\item 
\begin{equation}\label{propiedad-coercividad-sigma}
\Phi_0(|\b{x}|)=o\left(\int_{0}^{T}F(t,\b{x})\ dt\right)\;\;\mbox{as}\;\;|\b{x}|\to \infty.
\end{equation}
\end{enumerate}
Then  the action integral $I$ is coercive.
\end{thm}

\begin{proof}
By the decomposition $\b{u}=\overline{\b{u}}+\b{\tilde{u}}$,   Cauchy-Schwarz's inequality 
and \eqref{holder_cont-mu}, we have
\begin{equation}\label{cota-diferencia-F}
\begin{split}
&\left|\int_0^T F(t,\b{u})-F(t,\b{\overline{u}})\,dt\right|=
\left|\int_0^T \int_0^1 \nabla F(t,\b{\overline{u}}+s\b{\tilde{u}}(t))\ccdot \b{\tilde{u}}(t) \,ds \,dt\right|
\\
&\leq \int_0^T \int_0^1 b_1(t)\varphi_0(|\b{\overline{u}}+s\b{\tilde{u}}(t)|)|\b{\tilde{u}}(t)|\,ds\,dt+
\int_0^T \int_0^1 b_2(t)|\b{\tilde{u}}(t)|\,ds\,dt
\\
&=I_1+I_2.
\end{split}
\end{equation}
On the one hand, by H\"older's inequality and Sobolev's inequality, we estimate $I_2$ as follows
\begin{equation}\label{cota-i2}
I_2\leq \|b_2\|_{L^1} \|\b{\tilde{u}}\|_{L^{\infty}}\leq
C_1\|\b{\dot u}\orlnor.
\end{equation}
 where $C_1=C_1(\|b_2\|_{L^1}, T)$. 

On the other hand, since $\Phi_0 \in \Delta_2$ globally, then $\varphi_0 \in \Delta_2$ globally and 
consequently $\varphi_0$ is a quasi-subadditive function, i.e. there exists $C(\varphi_0)>0$ such that 
$\varphi_0(a+b)\leq C(\varphi_0)(\varphi_0(a)+\varphi_0(b))$ for every $a,b\geq 0$.
In this way, we have
\begin{equation}\label{pot-suma}
\varphi_0(|\b{\overline{u}}+s\b{\tilde{u}}(t)|)\leq
C(\varphi_0)[\varphi_0(|\b{\overline{u}}|)+\varphi_0(\|\b{\tilde{u}}\|_{L^{\infty}})].
\end{equation}
for every $s \in [0,1]$. 

Now,  inequality \eqref{pot-suma}, H\"older's inequality, Sobolev's inequality,
 the monotonicity, the subadditivity,  $\Delta_2$ condition of $\varphi_0$ imply that
\begin{equation}\label{cota-i1}
\begin{split}
I_1&
\leq C(\varphi_0)\bigg\{ \varphi_0(|\b{\overline{u}}|) \|b_1\|_{L^1} \|\b{\tilde{u}}\|_{L^{\infty}}+
 \|b_1\|_{L^1}\varphi_0(\|\b{\tilde{u}}\|_{L^\infty})\|\b{\tilde{u}}\|_{L^\infty}\bigg\}
\\
&\leq C_2 \bigg\{ \varphi_0(|\b{\overline{u}}|) \|\b{\dot{u}}\orlnor
+\varphi_0(\|\b{\dot u}\orlnor) \|\b{\dot u}\orlnor\bigg\}
%\\
%&=
%C_2 \bigg\{ f(|\b{\overline{u}}|) \|\b{\dot{u}}\orlnor
%+\varepsilon(\|\b{\dot{u}}\orlnor)\|\b{\dot{u}}\orlnor^{\alpha_{\Phi}}
%%f(\|\b{\dot u}\orlnor) \|\b{\dot u}\orlnor\bigg\}
\end{split}
\end{equation}
where $C_2=C_2(\varphi_0,T, \|b_1\|_{L^1} )$. 

Next, by Young's inequality with complementary functions $\Phi_0$ and $\Psi_0$ and the fact that 
$\Phi_0 \in \Delta_2$ globally, Young's equality \cite[Eq. 2.7-2.8]{KR}, \cite[Th. 3-(ii), p. 23]{rao1991theory}, we get
\begin{equation}\label{cota-i1-parcial}
 \begin{split}
\varphi_0(|\b{\overline{u}}|) \|\b{\dot{u}}\orlnor
&\leq 
\Psi_0(\varphi_0(|\b{\overline{u}}|))+
\Phi_0(\|\b{\dot{u}}\orlnor)
\\
&\leq 
|\b{\overline{u}}|\varphi_0(|\b{\overline{u}}|)
+\Phi_0(\|\b{\dot{u}}\orlnor)
\\
&\leq C(\Phi_0)
\Phi_0(|\b{\overline{u}}|)
+\Phi_0(\|\b{\dot{u}}\orlnor)
\end{split}
\end{equation}
and 
\begin{equation}\label{cota-i1-parcial-segunda}
\varphi_0(\|\b{\dot u}\orlnor) \|\b{\dot u}\orlnor
\leq 
C(\Phi_0) \Phi_0(\|\b{\dot u}\orlnor),
\end{equation}
with $C(\Phi_0)$ the constant that comes from the $\Delta_2$ condition on $\Phi_0$.

From \eqref{cota-i1}, \eqref{cota-i1-parcial}, \eqref{cota-i1-parcial-segunda} and \eqref{cota-i2},
%and the inequality $x^{r_1}\leq x^{r_2}+1$, for any $x\geq 0$ and $r_1\leq r_2$, 
we have
\begin{equation}\label{cota-i1-i2}
\begin{split}
I_1+I_2
&
\leq C_3
\bigg\{ 
\Phi_0(|\b{\overline{u}}|)
+\Phi_0(\|\b{\dot{u}}\orlnor)
+\|\b{\dot{u}}\orlnor
\bigg\}\\
&
\leq C_4
\bigg\{ 
\Phi_0(|\b{\overline{u}}|)
+\Phi_0(\|\b{\dot{u}}\orlnor)
+1
\bigg\}
\end{split}
\end{equation}
with $C_3$ and $C_4$ depending on $\Phi_0,T, \|b_1\|_{L^1}$ and $\|b_2\|_{L^1} $. The last inequality follows from the fact that $\Phi_0$ is an $N$-function, then there exists $C>0$ such that $\Phi_0(x)\geq Cx$ for every $x\geq 1$. Thus $x\leq C\Phi_0(x)+1$ for every $x\geq 0$.


In the subsequent estimates, we use  \eqref{cota_inf}, \eqref{cota-diferencia-F},
\eqref{cota-i1-i2}, the fact that $\Phi_0 \in \Delta_2$ and we get
\begin{equation}\label{cota_inf_I}
\begin{split}
I(\b{u})&\geq\alpha_0\rho_{\Phi}\left( \frac{\b{\dot{u}}}{\Lambda}\right)+\int_0^TF(t,\b{u})\ dt
\\ 
&=\alpha_0\rho_{\Phi}\left( \frac{\b{\dot{u}}}{\Lambda}\right)+ \int_0^T \left[F(t,\b{u})-F(t,\b{\overline{u}})\right]\ dt 
+  \int_0^TF(t,\b{\overline{u}})\ dt
\\
&\geq \alpha_0\rho_{\Phi}\left( \frac{\b{\dot{u}}}{\Lambda}\right)
-C_4 \Phi_0(\|\b{\dot u}\orlnor)
+\int_0^TF(t,\b{\overline{u}})\ dt-
C_4 \Phi_0(|\b{\overline{u}}|)-
C_4 
\\
&\geq
\alpha_0\rho_{\Phi}\left( \frac{\b{\dot{u}}}{\Lambda}\right)
-C_4 \Phi_0(\|\b{\dot u}\orlnor)
+H_{C_4, \Phi_0}(\b{\overline{u}})
-C_4 
\\&\geq
\alpha_0\rho_{\Phi}\left( \frac{\b{\dot{u}}}{\Lambda}\right)
-C_5 \Phi_0\left(\frac{\|\b{\dot u}\orlnor}{\Lambda} \right)
+H_{C_4, \Phi_0}(\b{\overline{u}})
-C_4 
\\&=
\alpha_0J_{C_6,\Phi_0}\left(\frac{\b{\dot u}}{\Lambda}\right)
+H_{C_4, \Phi_0}(\b{\overline{u}})
-C_4 
\end{split}
\end{equation}
where $C_5=C_5(\Phi_0,\Lambda,C_4)$ and $C_6=\frac{C_5}{\alpha_0}$.



Let $\b{u}_n$ be  a sequence in $\domi$ with 
$\|\b{u}_n\sobnor\to\infty$ and we have to prove that $I(\b{u}_n)\to\infty$. 
On the contrary, suppose  that for a subsequence, 
still denoted by $\b{u}_n$, $I(\b{u}_n)$ is upper bounded, i.e., there exists $M>0$ such that $|I(\b{u}_{n})|\leq M$. 
As $\|\b{u}_n\sobnor\to\infty$, from Lemma \ref{infinito-a-prom-upunto},  we have $|\overline{\b{u}}_n|+\|\b{\dot{u}}_n\orlnor\to \infty$. Passing to a subsequence, still denoted $\b{u}_n$, we can assume that $|\b{\overline u}_n|\to \infty$ or $\|\b{\dot{u}}_n\orlnor\to \infty$.
Now, Lemma \ref{lem_coer} implies that the functional $J_{C_6,\Phi_0}(\frac{\b{\dot u}}{\Lambda})$ is coercive,
and, by \eqref{propiedad-coercividad-sigma},
the functional $H_{C_4,\Phi_0}(\b{\overline{u}})$ is also coercive, then 
$J_{C_6,\Phi_0}(\frac{\b{\dot u}_n}{\Lambda}) \to \infty$ or $H_{C_4,\Phi_0}(\b{\overline{u}}_n)\to \infty$.
From \eqref{condA2}, we have that on a bounded set the functional $H_{C_4,\Phi_0}(\b{\overline{u}}_n)$ is lower bounded and also $J_{C_6,\Phi_0}(\frac{\b{\dot u}_n}{\Lambda})\geq 0$. 
Therefore,  $I(\b{u}_n)\to\infty$ as $\|\b{u}_n\sobnor\to\infty$ which contradicts the initial assumption on the behavior of $I(\b{u}_n)$. 
\end{proof}













% 
% 
% 
% 
% \subsection{versi\'on potencias por si acaso...} 
% \begin{thm}\label{coercitividad-r}
% Let  $\mathcal{L}$ be a lagrangian function satisfying \eqref{cotaL}, \eqref{cotaDxL}, \eqref{cotaDyL}, \eqref{cota_inf}  and $F$ satisfies condition (A). We assume the following conditions:
% \begin{enumerate}
% \item $\Psi\in\Delta_2$.
% \item There exist  non negative functions  $b_1,b_2 \in L^1_1$ and a constant $1<\mu<\alpha_{\Phi}$  such that 
% for any $\b{x}\in\rr^d$ and a.e. $t\in [0,T]$
% \begin{equation}\label{holder_cont-mu}
%   \left| \nabla F(t,\b{x}) \right|\leq b_1(t)|\b{x}|^{\mu-1}+b_2(t).
% \end{equation}
% \item There exists a real positive number $\sigma$ such that $\sigma>(\mu-1)\beta_{\Psi}$ and
% \begin{equation}\label{propiedad-coercividad-sigma}
% |\b{x}|^{\sigma}=o\left(\int_{0}^{T}F(t,\b{x})\ dt\right)\;\;\mbox{as}\;\;|\b{x}|\to \infty.
% \end{equation}
% \end{enumerate}
% Then  the action integral $I$ is coercive.
% \end{thm}
% 
% 
% 
% 
% \begin{proof} 
% By the decomposition $u=\overline{u}+\b{\tilde{u}}$,  Mean Value Theorem, Cauchy-Schwarz's inequality 
% and \eqref{holder_cont-mu}, we have
% \begin{equation}\label{cota-diferencia-F}
% \begin{split}
% &\left|\int_0^T F(t,\b{u})-F(t,\b{\overline{u}})\,dt\right|=
% \left|\int_0^T \int_0^1 \nabla F(t,\b{\overline{u}}+s\b{\tilde{u}}(t))\ccdot \b{\tilde{u}}(t) \,ds \,dt\right|
% \\
% &\leq \int_0^T \int_0^1 b_1(t)|\b{\overline{u}}+s\b{\tilde{u}}(t)|^{\mu-1}|\b{\tilde{u}}(t)|\,ds\,dt+
% \int_0^T \int_0^1 b_2(t)|\b{\tilde{u}}(t)|\,ds\,dt
% \\
% &=I_1+I_2.
% \end{split}
% \end{equation}
% On the one hand, by H\"older's inequality and Sobolev's inequality, we estimate $I_2$ as follows
% \begin{equation}\label{cota-i2}
% I_2\leq \|b_2\|_{L^1} \|\b{\tilde{u}}\|_{L^{\infty}}\leq
% C_1\|\b{\dot u}\orlnor.
% \end{equation}
%  where $C_1=C_1(\|b_2\|_{L^1}, T)$. On the other hand, as $s\in [0,1]$, we have
% \begin{equation}\label{pot-suma}
% |\b{\overline{u}}+s\b{\tilde{u}}(t)|^{\mu-1}\leq
% C(\mu)(|\b{\overline{u}}|^{\mu-1}+\|\b{\tilde{u}}\|_{L^{\infty}}^{\mu-1}).
% \end{equation}
% where $C(\mu)=2^{\mu-2}$, for $\mu\geq 2$ and $C(\mu)=1$, for $1<\mu<2$. Now,  inequality \eqref{pot-suma}, H\"older's inequality and Sobolev's inequality imply that
% \begin{equation}\label{cota-i1}
% \begin{split}
% I_1&\leq 
% C(\mu)\left(|\b{\overline{u}}|^{\mu-1} \int_0^T b_1(t) |\b{\tilde{u}}(t)|\,dt+
% \|\b{\tilde{u}}\|^{\mu-1}_{L^{\infty}} \int_0^T b_1(t)|\b{\tilde{u}}(t)| \,dt\right)
% \\
% &\leq C(\mu)\bigg\{ |\b{\overline{u}}|^{\mu-1} \|b_1\|_{L^1} \|\b{\tilde{u}}\|_{L^{\infty}}+
%  \|b_1\|_{L^1}\|\b{\tilde{u}}\|^{\mu}_{L^\infty}\bigg\}
% \\
% &\leq C_2 \bigg\{ |\b{\overline{u}}|^{\mu-1} \|\b{\dot{u}}\orlnor+ \|\b{\dot u}\orlnor^{\mu}\bigg\},
% \end{split}
% \end{equation}
% where $C_2=C_2(\mu,T, \|b_1\|_{L^1} )$. Let $\mu'$ be a positive constant such that $1<\mu\leq \mu'<\alpha_{\Phi}$. 
% Next, using Young's inequality with conjugate exponents $\mu'$ and $\frac{\mu'}{\mu'-1}$ 
%  we get
% \begin{equation}\label{cota-i1-parcial}
% |\b{\overline{u}}|^{\mu-1}   \|\b{\dot{u}}\orlnor
% \leq \frac{(\mu'-1)}{\mu'}|\b{\overline{u}|^{\sigma}}
% +\frac{1}{\mu'} \|\b{\dot{u}}\orlnor^{\mu'}
% \end{equation}
% where $\sigma=\frac{(\mu-1) \mu'}{\mu'-1}$. We point out that $\sigma$ is an arbitrary positive constant bigger than $(\mu-1)b_{\Psi}$.
% 
% From \eqref{cota-i1}, \eqref{cota-i1-parcial}, \eqref{cota-i2} and the inequality $x^{r_1}\leq x^{r_2}+1$, for any $x\geq 0$ and $r_1\leq r_2$, we have
% \begin{equation}\label{cota-i1-i2}
% \begin{split}
% I_1+I_2
% &\leq C_3\bigg\{ |\b{\overline{u}}|^{\sigma}
% + \|\b{\dot u}\orlnor^{\mu'}
% + \|\b{\dot u}\orlnor^{\mu}
% +\|\b{\dot u}\orlnor\bigg\}\\
% &\leq C_3\bigg\{ |\b{\overline{u}}|^{\sigma}
% + \|\b{\dot u}\orlnor^{\mu'}
% +1\bigg\}
% \end{split}
% \end{equation}
% with $C_3= C_3(\mu,T, \|b_1\|_{L^1},\mu' )$. In the subsequent estimates, we use the decomposition $u=\overline{u}+\b{\tilde{u}}$, \eqref{cota_inf}, \eqref{cota-diferencia-F},
% \eqref{cota-i1-i2} and we get
% \begin{equation}\label{cota_inf_I}
% \begin{split}
% I(\b{u})&\geq\alpha_0\rho_{\Phi}\left( \frac{\b{\dot{u}}}{\Lambda}\right)+\int_0^TF(t,\b{u})\ dt
% \\ 
% &=\alpha_0\rho_{\Phi}\left( \frac{\b{\dot{u}}}{\Lambda}\right)+ \int_0^T \left[F(t,\b{u})-F(t,\b{\overline{u}})\right]\ dt 
% +  \int_0^TF(t,\b{\overline{u}})\ dt
% \\
% &\geq \alpha_0\rho_{\Phi}\left( \frac{\b{\dot{u}}}{\Lambda}\right)
% -C_3 \|\b{\dot u}\orlnor^{\mu'}
% +\int_0^TF(t,\b{\overline{u}})\ dt-
% C_3 |\b{\overline{u}}|^{\sigma}-C_3\\
% &=\alpha_0J_{C_4,\mu'}\left(\frac{\b{\dot u}}{\Lambda}\right)
% + H_{C_3,\sigma}(\b{\overline{u}})-C_3,
% \end{split}
% \end{equation}
% where $C_4=\Lambda^{\mu'}C_3/\alpha_0$.
% 
% 
% Let $\b{u}_n$ be  a sequence in $\domi$ with 
% $\|\b{u}_n\sobnor\to\infty$ and we have to prove that $I(\b{u}_n)\to\infty$. 
% On the contrary, suppose  that for a subsequence, 
% still denoted by $\b{u}_n$, $I(\b{u}_n)$ is upper bounded, i.e., there exists $M>0$ such that $|I(\b{u}_{n})|\leq M$. 
% As $\|\b{u}_n\sobnor\to\infty$, from Lemma \ref{infinito-a-prom-upunto},  we have $|\overline{\b{u}}_n|+\|\b{\dot{u}}_n\orlnor\to \infty$.
% Then, there exists a subsequence of $\{\b{u}_n\}$, still denoted by $\b{u}_n$, which is not bounded.
% Then, 
% $|\b{\overline u}_n|\to \infty$ or $\|\b{\dot{u}}_n\orlnor\to \infty$.
% Now, Lemma \ref{lem_coer} implies that the functional $J_{C_4,\mu'}(\frac{\b{\dot u}}{\Lambda})$ is coercive,
% and, by \eqref{propiedad-coercividad-sigma},
% the functional $H(\b{\overline{u}})$ is also coercive, then 
% $J_{C_4,\mu'}(\frac{\b{\dot u}_n}{\Lambda}) \to \infty$ or $H(\b{\overline{u}}_n)\to \infty$.
% From \eqref{condA2}, we have that on a bounded set the functional $H(\b{\overline{u}}_n)$ is lower bounded; and, $J_{C_4,\mu'}(\frac{\b{\dot u}_n}{\Lambda})$ is also lower bounded  because the modular $\rho_{\Phi}\left(\frac{\b{\dot u}}{\Lambda}\right)$ is always bigger than zero. 
% Therefore,  $I(\b{u}_n)\to\infty$ as $\|\b{u}_n\sobnor\to\infty$ which contradicts the initial assumption on the behavior of $I(\b{u}_n)$. 
% \end{proof}
% 
% {\bf Leer y ver si es coherente lo anterior,  si conviene trabajar siempre con $u_n$ o habr\'ia que usar la notaci\'on de subsucesiones expl\'icita!!!}
% 
% {\bf Falta leer y corregir la  secci\'on que sigue del caso l\'imite!!!}

\section{Limit case $\mu=\alpha_{\Phi}$}

Coercivity was obtained even in the limit case $\mu=1$ and   $\mu=p$ 
(see \cite{zhao2004periodic, tang2010periodic})
assuming additional conditions on one of the $L^1$ 
functions that bound the nonlinearity.
These results lean on the fact that in $L^p$ spaces the norm and the modular coincides, 
that is, $\|\cdot\|_p^p=O(\int_0^T |\cdot|^p\,dt)$.
In Orlicz spaces, $\|\cdot\orlnor^\mu$ can be upper controlled by a modular provided that $\mu<\alpha_{\Phi}$ for any
$N$-function $\Phi$. 
Nevertheless,  the limit case does  not hold for any $\Phi$, i.e. in general $\|\cdot\orlnor^{\alpha_{\Phi}}=O(\int_0^T \Phi(|u|)\,dt)$ is false as can be seen as follows.


Let $\Phi$ be an $N$-function that satisfies the $\Delta_2$-condition. 
We claim that the inequality $\Phi(tu)\geq t^{\alpha_{\Phi}}\Phi(u)$ for any $u>0$ and for any $t\geq 1$ is false.

With the aim of proving the above assertion, we define the function 
\[\Phi(u)=
\left\{
\begin{array}{ll}
\frac{p-1}{p}u^p&u\leq e
\\
\frac{u^p}{\log u}-\frac{e^p}{p}&u>e
\end{array}
\right.\]
with $p>1$.

{\bf  Me parece que habr\'ia que decir algo que aclare los que se hace a continuaci\'on. 
Por ejemplo: 

Next, we will study the behaviour of $\Phi$/ establish some properties of $\Phi$/...}

\begin{thm}
If $p\geq \frac{1+\sqrt 2}{2}$, then $\Phi$ is an $N$-function.
\end{thm}


\begin{proof}
We have 
\[\Phi'(u)=\left\{
\begin{array}{cccc}
(p-1)u^{p-1}&:=&\varphi_1(u)& \mbox{if}\;u<e
\\
\frac{u^{p-1}}{\log u}(p-\frac{1}{\log u})&:=&\varphi_2(u)&\mbox{if}\; u>e
\end{array}
\right.
\]
and $\Phi$ is differentiable at $e$ because $\varphi_1(e)=\varphi_2(e)=(p-1)e^{p-1}$.

We will see that $\Phi'$ is increasing when $p\geq \frac{1+\sqrt {2}}{2}$.
Then, it is enough to see that $\varphi_1$ is increasing  on $[0,e]$ and $\varphi_2$ is increasing on 
$[e,\infty)$ for $p\geq \frac{1+\sqrt {2}}{2}$.

{\bf Tendr\'iamos que ver que $\varphi_1,\varphi_2$ son crecientes y que $\varphi_2\to \infty$ cuando
$u \to \infty????$ o basta con ver que $\varphi_2$ es creciente????}

$\varphi_1$ is an increasing function for $p>1$ and $\varphi_1(u)\to 0$ as $u \to 0$.

On the other hand, $\varphi_2(u)\to \infty$ as $u \to \infty$ provided that $p>1$.
And,  
$\varphi_2'(u)>0$
%\frac{u^{p-2}}{\log u} \left(p^2-p-\frac{2p}{\log u}+\frac{1}{\log u}+\frac{2}{\log^2 u}\right)\] 
on $[e,\infty)$ if and only if 
\begin{equation}\label{cuad-log-pos}
\left(p^2-p-\frac{2p}{\log u}+\frac{1}{\log u}+\frac{2}{\log^2 u}\right)>0.
\end{equation}
%If we take $\alpha:=\frac{1}{\log u}$, then we need 
%\[
%2\alpha^2+(1-2p)\alpha+(p^2-p)\geq 0
%\]
%which is true 
Now, as \eqref{cuad-log-pos} holds if and only if $p \notin(\frac{1-\sqrt2}{2},\frac{1+\sqrt2}{2})$, then $\varphi_2$ is an icreasing function when $p\geq \frac{1+\sqrt2}{2}$.
\end{proof}

{\bf  Conviene agregar m\'as cuentas en la prueba anterior, es decir, que la f\'ormula \eqref{cuad-log-pos} es una cuadr\'atica y bla....???}

\begin{thm}
There exists a constant $C>0$ such that 
\begin{equation}\label{cota-sup-indices}
\Phi(tu)\leq ct^p\Phi(u)\;\;t\geq 1, u>0.
\end{equation}
For every $\varepsilon>0$ there exists a constant $C=C(\varepsilon,p)$ such that
\begin{equation}\label{cota-inf-indices}
\Phi(tu)\geq Ct^{p-\varepsilon}\Phi(u)\;\;t\geq 1,u>0.
\end{equation}
\end{thm}

\begin{proof}
In order to prove \eqref{cota-sup-indices}, we analyze three cases.

If $u\leq tu\leq e$, then $\Phi(tu)=t^p\Phi(u)$ and \eqref{cota-sup-indices} holds with $C=1$.

If $u\leq e\leq tu$, as $\frac{e^p}{p}>0$  and $\log(tu)\geq 1$, we have 
$\Phi(tu)\leq t^pu^p= \frac{p}{p-1}t^p\Phi(u)$. Thus, \eqref{cota-sup-indices} holds with $C=\frac{p}{p-1}$.

If $e\leq u\leq tu$, then 
\begin{equation}\label{Phi-de-u-a-v}
\Phi(tu)\leq \frac{t^pu^p}{\log(tu)}\leq \frac{t^pu^p}{\log(u)}=\frac{pt^pv}{\log v}
\end{equation} where $v:=u^p$ and $v\geq e^p$.

If $\alpha>0$, the function $f(x)=\frac{x}{x-\alpha}$ is decreasing on $(\alpha,\infty)$.
And, the function $g(v)=\frac{pv}{\log v}$ is decreasing  on $[e^p,\infty)$. 
Therefore, $f \circ g$ is decreasing on $[e^p,\infty)$ and we have 
\[
(f\circ g)(v)=
\frac{\frac{pv}{\log v}}{\frac{pv}{\log v}-\frac{e^p}{p}}\leq 
e^p-\frac{e^p}{p}=\frac{p}{p-1}
\]
for every $v \geq e^p$.

In this way, from \eqref{Phi-de-u-a-v}, we have
\[
\Phi(tu)\leq \frac{pt^p}{p-1}\left(\frac{pv}{\log v}-\frac{e^p}{p}\right)=
 \frac{pt^p}{p-1}\left(\frac{u^p}{\log u}-\frac{e^p}{p}\right)
\]
and \eqref{cota-sup-indices} holds with $C=\frac{p}{p-1}$.

Now, we will prove \eqref{cota-inf-indices}.

If $u\leq tu \leq e$, \eqref{cota-inf-indices} is immediate because $t^p \geq t^{p-\varepsilon}$ for every $t \geq 1$, $p>1$ and $\varepsilon$ {\bf sufficiently small????}

If $u\leq e\leq tu$, as $f(t)=\frac{t}{\log t}$ is increasing on $[e,\infty)$ then $f(t)\geq e$ for every $t\geq e$. 
{\bf Habr\'ia que mirar en $[e^p,\infty)$ para que $f(t)\geq \frac{e^p}{p}$???? O, como $f$ es creciente y 
$e^p\leq (tu)^p$ entonces
$f((tu)^p)\geq f(e^p)$???}
\\
Now, 
\[
\begin{split}
\Phi(tu)=\frac{p(tu)^p}{\log (tu)^p}-\frac{e^p}{p}=
\frac{(p-1)(tu)^p}{\log(tu)^p}+\frac{(tu)^p}{\log (tu)^p}-\frac{e^p}{p}\geq 
\\
\frac{p-1}{p}\frac{(tu)^p}{\log(tu)}\geq
\frac{p-1}{p}\frac{t^{\varepsilon}}{\log t+1}t^{p-\varepsilon}u^p.
\end{split}
\]
Since $f(t)=\frac{t^{\varepsilon}}{\log t+1}$ attains its 
 minimum value $\varepsilon e^{1-\varepsilon}$ at 
$e^{\frac{1-\varepsilon}{\varepsilon}}$, \; then  
\[
\Phi(tu)\geq \frac{p-1}{p}\varepsilon e^{1-\varepsilon}t^{p-\varepsilon}u^p.
\]
If $e\leq u\leq tu$, then 
\[
\Phi(tu)=\frac{p-1}{p}\frac{(tu)^p}{\log(tu)}
+\frac{1}{p}\frac{(tu)^p}{\log(tu)}-\frac{e^p}{p}
\geq
\frac{p-1}{p}\frac{(tu)^p}{\log(tu)}=
\frac{p-1}{p}
\frac{t^{\varepsilon} \log (u)^{\varepsilon}}{\log(t^{\varepsilon}u^{\varepsilon})}
\frac{t^{p-\varepsilon}u^p}{\log u}
\]
Let $f(s)=\frac{sA}{\log s+1}$ with $s\geq 1$ and $A\geq \varepsilon$.
Then, the function $f$ attains a minimum at $s=e^{1-A}$; but, 
$s$ has to be bigger than 1, then it is necessary that
$\varepsilon\leq A\leq 1$.
And, the minimum value is $f(e^{1-A})=Ae^{1-A}\geq \varepsilon$.
If $A\geq 1$, $f$ attains the minimum at $s=1???$ and $f(1)=1$. 
Then, $f\geq \varepsilon$ and  therefore
\[
\Phi(tu)\geq \frac{p-1}{p}\varepsilon \frac{t^{p-\varepsilon}u^p}{\log u}\geq
\frac{p-1}{p}\varepsilon t^{p-\varepsilon}\Phi(u).
\]
\end{proof}

{\bf La prueba anterior est\'a rara.}





\begin{comentario}
The inequality 
\[
\Phi(tu)\geq Ct^p\Phi(u)
\] 
is false for every $C$ because for every $u\geq e$ we have 
\[
\lim\limits_{t \to \infty}\frac{\Phi(tu)}{t^p\Phi(u)}=
\lim\limits_{t \to \infty} \frac{\frac{u^p}{\log(tu)}-
\frac{e^p}{pt^p}}{\frac{u^p}{\log u}-\frac{e^p}{p}}=0
\]
\end{comentario}



{\bf El l\'imite intermedio de la f\'ormula anterior se podr\'ia quitar.}


\begin{thm}
$\alpha_{\Phi}=\beta_{\Phi}=p$
\end{thm}

\begin{proof}
From \eqref{MO_indices} and \eqref{cota-sup-indices}, we get 
\[
\beta_{\Phi}=\lim\limits_{t \to \infty} \frac{\log\left[\sup\limits_{u>0} \frac{\Phi(tu)}{\Phi(u)}\right]}{\log t}
\leq
\lim \limits_{t \to \infty} \frac{\log C+p\log t}{\log t}=p.
\]
On the other hand, employing \eqref{MO_indices} and performing some elementary calculations, we obtain
\[
\alpha_{\Phi}=
\lim\limits_{t \to 0^+} \frac{\log\left[\sup\limits_{u>0} \frac{\Phi(tu)}{\Phi(u)}\right]}{\log t}=
\lim\limits_{s \to \infty} \frac{\log\left[\sup\limits_{v>0} \frac{\Phi(v)}{\Phi(sv)}\right]^{-1}}{\log s}=
\lim\limits_{s \to \infty} \frac{\log\left[\inf\limits_{v>0} \frac{\Phi(sv)}{\Phi(v)}\right]}{\log s}
\]
where $v:=tu$ and $s:=\frac{1}{t}$.
Then, using \eqref{cota-inf-indices},  for every $\varepsilon>0$ we have
\[
\alpha_{\Phi}=
\lim\limits_{s \to \infty} \frac{\log\left[\inf\limits_{v>0} \frac{\Phi(sv)}{\Phi(v)}\right]}{\log s}\geq
\lim\limits_{s \to \infty} \frac{\log C+(p-\varepsilon)\log s}{\log s}\geq p-\varepsilon,
\]
therefore $\alpha_{\Phi}\geq p$. 

Finally, as $\alpha_{\Phi}\leq \beta_{\Phi}\leq p$, we get  
$\alpha_{\Phi}=\beta_{\Phi}=p$.
\end{proof}



Now, we are able to see that 
\[
\rho_{\Phi}(u)=\int_0^T \Phi(|u|)\,dx\geq C\|u\orlnor^{\alpha_{\Phi}}=C\|u\orlnor^p
\]
is false.

In fact, if we take $u\equiv t>0$, then $\|u\orlnor^p=C_1t^p$ where $C_1=\|1\orlnor$ and
$\int_0^T \Phi(|u|)\,dx=C_2\Phi(t)$ with $C_2=T$. 
Then, if $\rho_{\Phi}(u)\geq C\|u\orlnor^p$ were true, then $\Phi(t)\geq C t^p$ would also be true; however, this
last inequality is false.
 



\section*{Acknowledgments}
The authors are partially supported by a UNRC grant number 18/C417. The first author is  partially supported by a  UNSL grant number 22/F223. 


 \bibliographystyle{elsarticle-num} 
 \bibliography{biblio}


\end{document}
